---
title: "Olympics2016"
author: "Deepal Singh"
date: "January 11, 2019"
output: 
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Curation and Parsing

Data Curation is an important step during the data analysis process. It provides meaningful access to the data.It involves constructing a structure that sorts the data into a meaningful retrieval system and gives us intellectual control over content.

### Importing the libraries and reading table from url

```{r reading_table}
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(naniar)
library(measurements)


olympics_data <- read.csv(file = "C:\\Users\\Mandevi\\Documents\\athletes.csv", na.strings = c("", " ", "NA"), stringsAsFactors = FALSE)

as_data_frame(olympics_data)

```

### Tidying the data

This section includes data preparation, that is taking raw datasets and make changes to turn messy data into tidy data. If we open the url and see the table, we can observe that there are lots of things that can be made simple. For instance,I have converted date of birth to age and also renamed variables for readability.


```{r tidying_data}

library(eeptools)
library(lubridate)


# Removing id because we don't need it
tidy_olympics<- subset(olympics_data, select = -c(id))%>%
  na.omit(olympics_data)

# meter to feet and inches function
ft_and_inches <- conv_unit(tidy_olympics$height, "m", "ft")
tidy_olympics$height<- round(ft_and_inches, digits = 2)

#  changing dob to age
conv_age <- function(dob, age.day = Sys.Date(), units = "years", floor = TRUE) {
    calc.age = interval(dob, age.day) / duration(num = 1, units = units)
    if (floor) 
      return(as.integer(floor(calc.age)))
    return(calc.age)
}

tidy_olympics$dob <- as.Date(tidy_olympics$dob, format = "%m/%d/%y")
tidy_olympics$dob <- conv_age(tidy_olympics$dob)

# renaming the nationality into Country_code
names(tidy_olympics)[names(tidy_olympics)== "nationality"] <-"Country_Code"
names(tidy_olympics)[names(tidy_olympics)== "dob"] <-"age"
names(tidy_olympics)[names(tidy_olympics)== "name"] <-"Athlete_name"

as_data_frame(tidy_olympics)

```

## Exploratory Data Analysis
Exploratory Data Analysis helps to better understand the tidy data, and help us make decisions about appropriate statistical or Machine Learning methods, or data transformations.

### Number of medals won by country

This table is on the country winning higher number of medals and USA has highest wins.

```{r medal_analysis}
tidy_olympics$Total_medals = rowSums(tidy_olympics[,8:10])

olympics_data_subset<- tidy_olympics%>%
  filter(Total_medals != 0)%>%
  select("Athlete_name","Country_Code","sex", "sport", "Total_medals")%>%
  arrange(Country_Code)
  

country <- olympics_data_subset%>% 
  group_by(Country_Code)%>%
  summarise(total_wins = sum(Total_medals))%>%
  arrange(desc(total_wins))
country

```

### Popular sport according to the numbers participated

From the table below, it can be verified that many particpants take part more in aquatics.

```{r popular sport}
popular_sport<- olympics_data_subset%>% 
  select(Country_Code, sport)%>%
  group_by(sport)%>%
  summarize(participants_num = length(sport))%>%
  arrange(desc(participants_num))

popular_sport

```

### Plot of the frequency of participants playing particular sport

```{r plot_frequency}

popular_sport%>% 
  ggplot(aes(y=participants_num, x = sport, group = 1))+ 
  geom_line(color= "red", size= 1.5)+
  xlab("Sport") +   
  ylab("Frequency ofsports") +   
  ggtitle("Frequency of partipants playing particular sport")+
   theme(axis.text.x = element_text(angle=90, hjust=0.5))

```

### Analyzing males vs females

This is the barplot to see if there is difference between males and females involved in sports. And it looks like there is not much difference. This is how my hypothesis was developed for further analysis.

```{r male_vs_female}
gender_data<- olympics_data_subset%>%
  group_by(sex)%>%
  summarise(Gender_Count = length(sex))%>%
  ggplot(aes(x=sex, y = Gender_Count))+
  geom_bar(stat="identity", mapping = aes(fill= sex, color=sex))+
  xlab("Gender") +   
  ylab("Number of participants") +   
  ggtitle("Number of participants based on Gender")
gender_data
  
```

## Hypothesis Testing & Machine learning
I will be performing a statistical  hypothetical testing. Hypothesis testing is, in a way, the formal way of validating the hypothesis claim that is made about a population. Statistical Hypothesis can be classified into into 2 types; null hypothesis and alternative hypothesis.Through EDA, I was able find that there is not much Gender differences for involvement in sports. The common stereotype is that men are most capable in physical activites and women would lag behind.So, I developed the hypothesis to test if this stereotype is really true.

### Null Hypothesis: 
There is no difference bewtween males and females involved in # sports in olympics.

### Alternative Hypothesis: 
There are differences between males and females involved in sports in olympics

### T_test

A t-test looks at the t-statistic, the t-distribution and degrees of freedom to determine the probability of difference between variables. I am performing a t_test with 95% confidence interval and it is going to be two sided. Confidence interval is the range of values defined such that there is specified probability that the value of parameter lies within.

```{r t_test}
g_data <- olympics_data_subset%>%
  select(sport, sex)%>%
  group_by(sport, sex)%>%
  summarize(g_count = length(sex))

 
male <- g_data%>%filter(sex== "male")
  

female <- g_data%>%filter(sex == "female")

t.test(male$g_count, female$g_count, mu = 0, con4f.level = 0.95, alternative = "two.sided", paired = FALSE)

```

The table shows t = 0.32997, with the degree of freedom df = 51.724. The p-value equals .7428. With this given p-value and using significance level of 0.05, we fail to reject the null hypothesis. BUt we reject the alternative hypothesis saying that there is a difference between the male and the females. Therefore, we cannot conclude that there is a significant difference between the physical qualities and sport involvement of female and male players.


A 95% confidence interval for the difference in mean count is given by
```{r t_test2}
difference <- male$g_count - female$g_count
t.test(difference)
```
From the above data, the confidence interval is between -1.670905 and 6.633868. This shows that 95% of the true population mean differences between these two  gender are in this intervals.

## Machine learning

### linear regression
One of the most used techniques is linear regression where we can examine the potential relationship between variable and a set of one or more variables. Therefore, I will examine linear model for gender count.Following code uses lm function to interpret the data.


```{r linear_model}
library(broom)

linear_model_of_gender <- lm(g_count~sex, data = g_data, family = binomial)%>%
  tidy()
linear_model_of_gender
```

According to this model, when the male is not counted, the total number of women counterparts would be 31.222. 

### Regresssion tree graph

Following is the plot which gets a violin plot for the count of gender.

```{r d_tree}

plot_tree<-g_data %>%
  ggplot(aes(y = g_count, x= sex)) +
  geom_violin(aes(fill= sex, color=sex)) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle=90, hjust=0.5))
  
plot_tree

```


### Regression for sport and G_count

I am going to consider another linear regresion model which considers Sport and G_count.
The data distributed is unsymmetric and boxing seem to have high number of residuals.

```{r two}
regression <- lm(g_count~sport,data=g_data)%>% 
  augment()
regression%>% 
  ggplot(aes(x=factor(sport), y=.resid)) + 
  geom_boxplot() +
  labs(title="Residuals across Sport over two sexes", x = "sport", y = "residuals")+
  theme(axis.text.x = element_text(angle=90, hjust=0.5))
```



#Plot

```{r line_graph}
g_data%>%
  ggplot(aes(y= g_count, x = sport, color = sex, group = 1))+ 
  geom_point()+ 
  theme(axis.text.x = element_text(angle=90, hjust=0.5))
```

Based on tha above plot, it seems considering the sport would be beneficial to determine the gender involvement in the specified sport.We can observe that female are more involved in many sports than male. But male are ranked higher than female. Therefore, due to women's involvement in lot of sports than men, it makes it look that there is not significant difference between male and female counter parts.

